An Efficient Interpolation Technique for Jump Proposals in Reversible-Jump Markov Chain Monte Carlo Calculations

### 2011-01-06 paper.tex
Capitalized Monte Carlo in abstract.

\begin{abstract}
  Selection among alternative theoretical models given an observed data set is an important challenge in many areas of physics and astronomy, ranging from gravitational-wave data analysis to measurements of black-hole masses in X-ray binaries.  Reversible-jump Markov chain
  monte carlo (RJMCMC) is an extremely powerful technique for
  performing Bayesian model selection, but it suffers from a
  fundamental difficulty: it requires jumps between model parameter
  spaces, but cannot retain a memory of the favored locations in more
  than one parameter space at a time.  Thus, a naive jump between
  parameter spaces is unlikely to be accepted in the MCMC algorithm
  and convergence is correspondingly slow.  Here we demonstrate an
  interpolation technique that uses samples from single-model MCMCs to
  propose inter-model jumps from an approximation to the single-model
  posterior of the target parameter space.  The interpolation
  technique, based on a kD-tree data structure, is adaptive and
  efficient in arbitrary dimensions.  We show that our technique leads
  to dramatically improved convergence over naive jumps in an RJMCMC,
  and compare it to other proposals in the literature to improve the
  convergence of RJMCMCs.  We also discuss the use of the same
  interpolation technique in two other contexts: as a convergence test
  for a single-model MCMC and as a way to construct efficient
  ``global'' proposal distributions for single-model MCMCs without
  prior knowledge of the structure of the posterior distribution.
\end{abstract}

>>>>

\begin{abstract}
  Selection among alternative theoretical models given an observed data set is an important challenge in many areas of physics and astronomy, ranging from gravitational-wave data analysis to measurements of black-hole masses in X-ray binaries.  Reversible-jump Markov chain
  Monte Carlo (RJMCMC) is an extremely powerful technique for
  performing Bayesian model selection, but it suffers from a
  fundamental difficulty: it requires jumps between model parameter
  spaces, but cannot retain a memory of the favored locations in more
  than one parameter space at a time.  Thus, a naive jump between
  parameter spaces is unlikely to be accepted in the MCMC algorithm
  and convergence is correspondingly slow.  Here we demonstrate an
  interpolation technique that uses samples from single-model MCMCs to
  propose inter-model jumps from an approximation to the single-model
  posterior of the target parameter space.  The interpolation
  technique, based on a kD-tree data structure, is adaptive and
  efficient in arbitrary dimensions.  We show that our technique leads
  to dramatically improved convergence over naive jumps in an RJMCMC,
  and compare it to other proposals in the literature to improve the
  convergence of RJMCMCs.  We also discuss the use of the same
  interpolation technique in two other contexts: as a convergence test
  for a single-model MCMC and as a way to construct efficient
  ``global'' proposal distributions for single-model MCMCs without
  prior knowledge of the structure of the posterior distribution.
\end{abstract}
